### 2021-10-11 chapters/acknowledgements.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}
\chapter{Acknowledgements}

The most important person for me to acknowledge is my wife Andrea. 
It was only with her forbearance and encouragement that I was able to return to University, obtain my Honours in Advanced Science (Research) and continue on towards my doctorate over a total of a seven year period.
I would also like to thank my daughters: Lauren, Sophie and Phoebe, who have all literally grown up during this period.
I look forward to spending more time with you after this submission.
And of course, I am pleased to acknowledge my supervisors, Paul and Andy, who in their infinite patience encouraged me on this journey, particularly when covid-19 tried to stop the world in its tracks.
My peers in PhD room 125 both present and past, have made life interesting and fun over the last three years.
I'm sure you are all heading towards brilliant futures.
This research was supported by the Faculty of Science Dean's Postgraduate Research Scholarship without which I would not have had the opportunity to attempt this.

\end{document}