### 2021-10-11 chapters/copyright.tex

\documentclass[../Thesis.tex]{subfiles}

\begin{document}

\noindent This thesis must be used only under the normal conditions of “fair dealing” under the Copyright Act. It should not be copied or closely paraphrased in whole or in part without the written consent of the author. Proper written acknowledgement should be made for any assistance obtained from this thesis. I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner’s permission.
\\[1cm]
\copyright\ \  Paul Easter

\end{document}