### 2021-10-11 chapters/declaration.tex

\documentclass[../Thesis.tex]{subfiles}

\begin{document}
\chapter{Declaration of Authorship}
I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes two original papers published in peer reviewed journals and one submitted publications. The core theme of this thesis is probing the physics of binary neutron star post-merger remnants with gravitational waves. The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the School of Physics and Astronomy under the supervision of Assoc. Prof. Paul Lasky and Dr. Andrew Casey.

The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.

I have renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

In the case of Chapters 2, 3,and 4 my contribution to the work is shown on the following page. \\[0.75cm]
%
\tabto{0cm}\textbf{Student name:}  Paul John Easter \\[0.75cm]
\tabto{0cm}\textbf{Student signature:} \tabto{10cm} Date: 8 August 2021 \\[0.75cm]
%
%
I hereby certify that the above declaration correctly reflects the nature and extent of the student’s and co-authors’ contributions to this work. In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors. \\[0.75cm] 
%
\tabto{0cm}\textbf{Main Supervisor name:} Paul D Lasky \\[0.75cm]
\tabto{0cm}\textbf{Main Supervisor signature:}	\tabto{10cm} Date: 8 August 2021

\pagebreak

\begin{turn}{-90}

% \begin{table}
\begin{tabular}{|p{0.12\linewidth}|p{0.40\linewidth}|p{0.15\linewidth}|p{0.3\linewidth}|p{0.5\linewidth}|p{0.15\linewidth}|}
\hline
Thesis Chapter	& Publication Title	& Status	 
& Student contribution & Co-author contribution	& Monash student co-author\\ \hline \hline

2	
& Computing Fast and Reliable Gravitational Waveforms of Binary Neutron Star Merger Remnants	
& Published (PRD)	
& 75\% \newline Concept, method, code development, analysis, writing	 
& P. Lasky/A. Casey \newline 15\%-Concept, method, analysis, writing \newline L. Rezzolla/K. Takami \newline 10\% - analysis, writing 
& No\\
\hline
3	
& Detection and parameter estimation of binary neutron star merger remnants	
& Published (PRD)	
& 70\% \newline Concept, method, code development, analysis, writing	 
& S. Ghonge \newline 10\%-Concept, method, code development, analysis, writing \newline 
F. Hernandez Vivanco \newline 10\% - Concept, method, code development, analysis, writing \newline 
P. Lasky/A. Casey \newline 5\%-Concept, method, analysis, writing \newline 
J. Clarke/K. Chatziioannou \newline 5\% - Concept, analysis, writing
& Yes \newline Hernandez Vivanco\\
\hline

4	& Can we measure the collapse time of a post-merger remnant for a future GW170817-like event? 
& Submitted (PRD)
& 75\%  Concept, method, code development, analysis, writing	 
& P. Lasky/A. Casey \newline 25\%-Concept, method, analysis, writing 
& No\\
\hline

\hline				
\end{tabular}
% \end{table}

\end{turn}

\end{document}